---
title: "Eating Out"
author: "Ashley Cook"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "June 26 2023"    
    
---



## There were two years in which the General Social Survey asked about the frequency of eating out.
### Find those years and for the first year the question was asked download that years data in a stata format - its name will be "GSSxxxx.dta" where the xxxx is replaced by the year.  For example if the year was 1983 its name would be GSS1983.dta. Your assignment is to download the GSS yearly dataset for the first year the question was asked and then  write up the answer to the question by editing the  code block in line 31 to eval=TRUE and finish the job. 

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


##methods
  
Today, we will be reading and showing results of different races that eat out and eat in. We will be adding a ggplot, numerical, inferential stats, tables and fisher stats that show the p value and percentages of the results. Below I have various of different codes that show which race has the highest percentage of eating out, the odds and the lowest percentage of eating out too. 

```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
#require(haven)
require(foreign)
```


```{r, eval=TRUE, warning=FALSE}
#GSS <-(read_dta(file="GSS1998.dta))
GSS<- readstata13::read.dta13(file="GSS1998.dta")
GSS <-droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <-na.omit(mydata)

```


```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata, type="percent")

```
  
In the percent wise there was not much of a difference in th percent of eating out from black and other races. The white people were eating out more than the other races. 



```{r, eval=FALSE}
GSS <-read_data(file="GSS1998.dta")
GSS <- readstata13::read.dta13(file= "GsS1998.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```


```{r}
dd2 <- mydata %>% group_by(race,eatout) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=eatout))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=eatout)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The one on the left is by count and the one on th right is by percent. 




```{r}
table1 <- xtabs(~race + eatout, data=mydata)
rowPerc(table1)
colPerc(table1)
```
  
We have a numerical results of white people who are eating out the most and almost 70 percent of black people eat out, while 28 of them eat in. There is more white people out of the races, but they have the least amount of people who eat in. 

```{r}
chisq.test(table1)
chisqtestGC(table1)
```
  
There is a really low P-value from this test of .000002287. 

```{r}
fisherXY(mydata$race,mydata$eatout)
```
  
945 white people who ate out and 192 that did not. While the other race there was 75 that ate out and 29 did not. The odds of a white person is 2.2 times higher than a black person eating out. the white people almost 2 times greater than the black people do by 1.9 times often than the other race does. 

```{r}
fisherXY(mydata$race,mydata$eatout,printAll = TRUE)
```
  
The odds of the two races eating out where 0 is between the odds and the odds are .06 and 1 is in the middle. Where 1 being the middle, they are both different, meaning the black and white people eating out and in. The P value of white people is greater than 0.05, while the p-value is lower for black. The white and black have a difference of 2.2 eating out. Fisherxy is the analog to Tuki HD. 

##Conclusion

Overall, there was a lot more white people than the other races and they had a greater p-value and percentage of them eating out but they had the less amount of the other races of them eating in. The black people had the second highest amount of people eating out, but had the most of people eating in. Above the conclusion, there was graphs, charts and data that shows the amounts and percentages of how many people eat out and in. 